---
title: "GARP analysis on glioblastoma (GBM)"
date: "`r Sys.Date()`"
author: Yuzhou Chang
output:
  rmdformats::readthedown:
    self_contained: true
    thumbnails: true
    lightbox: true
    gallery: false
    highlight: tango
---

library(rmarkdown)
```{r setup, include=FALSE,message=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(ggpubr)
library(survival)
library(lubridate)
library(gtsummary)
library(tidycmprsk)
library(survminer)
library(ggplot2)
library(grid)
library(gridExtra)
```

# Data Overview
The Data (TCGA_cell_2013) have 152 patients and the expression data was downloaded from [cBioPortal](!https://www.cbioportal.org/datasets) and the original paper is [here](!https://pubmed.ncbi.nlm.nih.gov/24120142/)

## TCGA_cell_2013 information

This data contains 152 patients, including 39 classical, 49 Mesenchymal, 26 Neural, 29 Proneural, and 9 others ( G-CIMP). As you care about four subtypes, I exclude other categories and only keep 143 patients with four subtype categories. 

```{r,echo = F}
load("/bmbl_data/yuzhou/collaborative/Li_lab/Dan/analysis/TCGA_cell_2013_meta.Rdata")
TCGA_cell_2013_meta$censor <- ifelse(TCGA_cell_2013_meta$OS_STATUS == "1:DECEASED",1,0)
# new.meta <- new.meta[new.meta$survival_time != "NA",]
TCGA_cell_2013_meta[,c(2,3)] %>% tbl_summary()
TCGA_cell_2013_meta <- TCGA_cell_2013_meta[TCGA_cell_2013_meta$tutmor_type != "G-CIMP" & !is.na(TCGA_cell_2013_meta$tutmor_type),]

```


# Effect of higher GARP expression (i.e. top 10% vs bottom 10%) on survival of GBM patients.
## TCGA_cell_2013 survival analysis for all tumor subtypes
```{r, echo = F}
TCGA_cell_2013_meta$GARP_binary <- "no_group"
TCGA_cell_2013_meta$GARP_binary <- ifelse(TCGA_cell_2013_meta$GARP_expression > quantile(TCGA_cell_2013_meta$GARP_expression,0.9,na.rm = T ),"high_expression",TCGA_cell_2013_meta$GARP_binary)
TCGA_cell_2013_meta$GARP_binary <- ifelse(TCGA_cell_2013_meta$GARP_expression < quantile(TCGA_cell_2013_meta$GARP_expression,0.1,na.rm = T),"low_expression",TCGA_cell_2013_meta$GARP_binary)
# table(TCGA_cell_2013_meta$GARP_binary )
sub.meta.GARP <- TCGA_cell_2013_meta[TCGA_cell_2013_meta$GARP_binary!="no_group",]
sub.meta.GARP$OS_time <- as.numeric(sub.meta.GARP$OS_time)
km_trt_fit <- survfit(Surv(OS_time, censor) ~ GARP_binary, data=sub.meta.GARP)
ggsurvplot(km_trt_fit,pval = T)

```

## TCGA_cell_2013 survival analysis for four tumor subtypes 

### Survival analysis for classical
```{r, echo =F}
TCGA_cell_2013_meta_Classical <- TCGA_cell_2013_meta[TCGA_cell_2013_meta$tutmor_type == "Classical",]
sub.meta_Classical.GARP <- TCGA_cell_2013_meta_Classical[TCGA_cell_2013_meta_Classical$GARP_binary!="no_group",]
sub.meta_Classical.GARP$OS_time <- as.numeric(sub.meta_Classical.GARP$OS_time)
km_trt_fit <- survfit(Surv(OS_time, censor) ~ GARP_binary, data=sub.meta_Classical.GARP)
ggsurvplot(km_trt_fit,pval = T) + labs(title = "classical") 
```

### Survival analysis for mesenchymal  
```{r, echo = F}
TCGA_cell_2013_meta_Mesenchymal <- TCGA_cell_2013_meta[TCGA_cell_2013_meta$tutmor_type == "Mesenchymal",]
sub.meta_Mesenchymal.GARP <- TCGA_cell_2013_meta_Mesenchymal[TCGA_cell_2013_meta_Mesenchymal$GARP_binary!="no_group",]
sub.meta_Mesenchymal.GARP$OS_time <- as.numeric(sub.meta_Mesenchymal.GARP$OS_time)
km_trt_fit <- survfit(Surv(OS_time, censor) ~ GARP_binary, data=sub.meta_Mesenchymal.GARP)
ggsurvplot(km_trt_fit,pval = T,palette = c("#ff8080", "#5757f9"), ggtheme = theme_classic())+ labs(title = "Mesenchymal")
```

### Survival analysis for  neural  
```{r, echo = F}
TCGA_cell_2013_meta_Neural <- TCGA_cell_2013_meta[TCGA_cell_2013_meta$tutmor_type == "Neural",]
sub.meta_Neural.GARP <- TCGA_cell_2013_meta_Neural[TCGA_cell_2013_meta_Neural$GARP_binary!="no_group",]
sub.meta_Neural.GARP$OS_time <- as.numeric(sub.meta_Neural.GARP$OS_time)
km_trt_fit <- survfit(Surv(OS_time, censor) ~ GARP_binary, data=sub.meta_Neural.GARP)
ggsurvplot(km_trt_fit,pval = T)+ labs(title = "Neural") 
```

### Survival analysis proneural  
```{r, echo = F}
TCGA_cell_2013_meta_Proneural <- TCGA_cell_2013_meta[TCGA_cell_2013_meta$tutmor_type == "Proneural",]
sub.meta_Proneural.GARP <- TCGA_cell_2013_meta_Proneural[TCGA_cell_2013_meta_Proneural$GARP_binary!="no_group",]
sub.meta_Proneural.GARP$OS_time <- as.numeric(sub.meta_Proneural.GARP$OS_time)
km_trt_fit <- survfit(Surv(OS_time, censor) ~ GARP_binary, data=sub.meta_Proneural.GARP)
ggsurvplot(km_trt_fit,pval = T)+ labs(title = "Proneural") 

```


# Effect of GARP expression on the phenotype of GBM
In this analysis, we will compare GARP expression and GARP pathway activity among different GBM subtypes. The pathway activity is calculated by the GSVA algorithm.

## TCGA_cell_2013 

We calculate p-values for any two groups regarding GARP expression and GARP pathway activity. 

* Classical vs Mesenchymal 
* Classical vs Proneural
* Classical vs Neural
* Mesenchymal vs Proneural
* Mesenchymal vs Neural
* Proneural vs Neural

###  GARP expression

```{r, echo = F,warning=F}
TCGA_cell_2013_meta_GARP <-TCGA_cell_2013_meta[,c("tutmor_type","GARP_expression")]
TCGA_cell_2013_meta_GARP <- TCGA_cell_2013_meta_GARP[!is.na(TCGA_cell_2013_meta_GARP$GARP_expression),]
my.comparison <- list(c("Classical","Mesenchymal"),
                      c("Classical", "Proneural"),
                      c("Classical","Neural"),
                      c("Mesenchymal","Proneural"),
                      c("Mesenchymal","Neural"),
                      c("Neural","Proneural"))



ggboxplot(TCGA_cell_2013_meta_GARP, x = "tutmor_type", y = "GARP_expression",
          color = "tutmor_type", palette = "jco")+ 
  stat_compare_means(comparisons = my.comparison)

```

### GARP pathway activity

```{r, echo = F,warning=F}
TCGA_cell_2013_meta_GARP <-TCGA_cell_2013_meta[,c("tutmor_type","GARP_geneset")]
TCGA_cell_2013_meta_GARP <- TCGA_cell_2013_meta_GARP[!is.na(TCGA_cell_2013_meta_GARP$GARP_geneset),]
my.comparison <- list(c("Classical","Mesenchymal"),
                      c("Classical", "Proneural"),
                      c("Classical","Neural"),
                      c("Mesenchymal","Proneural"),
                      c("Mesenchymal","Neural"),
                      c("Neural","Proneural"))



ggboxplot(TCGA_cell_2013_meta_GARP, x = "tutmor_type", y = "GARP_geneset",
          color = "tutmor_type", palette = "jco")+ 
  stat_compare_means(comparisons = my.comparison)

```

